---
title: "Lab 3"
author: "Kyle Payne"
date: "September 9, 2015"
output: html_document
runtime: shiny
---


# Parameters
A parameter is a numerical characteristic of a population. Think, average overall rate of calls to a call center, or average height in a goat population, or the true variance of evoked potential response in a study. The parameter does not depend upon the sample for its calculation. 

# Statistic
A statistic is a numerical quantity calculated from a sample. These are sample means 

$$ \bar{X} = \frac{1}{n}\sum_{i=1}^{n}X_i $$ 

Sample Variance

$$ S^2 = \frac{1}{n-1}\sum_{i=1}^{n}(X_i - \bar{X})^2 $$

and so on ...

# Binomial Plot
Choose the parameters
```{r echo = FALSE}
sliderInput("bin_param",
             h1("Binomial Parameter"),
             min=0, max=1,
              value=0.5)
numericInput("sample_size",
             h1("Sample Size"),
             value=10, 
             min=0,
             max=10000)
```



```{r, echo=FALSE}
  renderPlot({
    print(input$bin_param)
    print(input$sample_size)

    p <- input$bin_param
    n <- input$sample_size
    s <- 10000
    rands <- rbinom(s, n, prob=p)
    print(rands)
    hist(rands, prob=T)
    
    norm_mean <- n*p
    norm_sd <- sqrt(n*p*(1-p))
    
    r_norms <- rnorm(s, mean=norm_mean, sd=norm_sd)
    lines(density(r_norms))
  })
```

# Normal
```{r echo = FALSE}
sliderInput("mean",
             h1("Mean"),
             min=-50, max=50,
              value=0)
numericInput("variance",
             h1("Variance"),
             value=1, 
             min=0,
             max=10000)
```



```{r, echo=FALSE}
  renderPlot({
    print(input$bin_param)
    print(input$sample_size)

    mean <- input$mean
    n_var <- input$variance
    s <- 10000
    rands <- rnorm(s, mean=mean, sd=n_var)
    plot(density(rands))
    abline(v=mean, lty=2, lwd=4, col="skyblue")
    abline(v = mean + sqrt(n_var), lty=2, lwd=4, col="red")
    abline(v = mean - sqrt(n_var), lty=2, lwd=4, col="red")
  })
```



```{r, echo=FALSE}

numericInput("num_flips",
             h1("Number of Flips"),
             value=10, 
             min=0,
             max=10000)
sliderInput("thresh",
             h1("Probability of Head"),
             value=0.5,
             min=0, 
             max=1)
```


```{r, echo=FALSE}
library(ggplot2)

renderPlot({
#Initialze the variables
  
sequenceOfCoinTosses <- sample(c(0,1), input$num_flips,
                               prob=c(1-input$thresh, input$thresh),
                               replace = TRUE)

cum_sum <- cumsum(sequenceOfCoinTosses)/seq_along(sequenceOfCoinTosses)

df <- data.frame(trials = 1:input$num_flips, relHH = cum_sum)


ggplot(aes(x=trials, y=relHH), data=df) + geom_line()
})
```


